---
title: "Install"
format:
  html:
    toc: true
execute:
  eval: false
---

The vetiver framework for MLOps tasks is built for data science teams using R and/or Python, with a native, fluent experience for both. It is built to be extensible, with methods that can support many kinds of models.

::: panel-tabset

## Python

You can use vetiver with:

- [scikit-learn](https://scikit-learn.org/)
- [PyTorch](https://pytorch.org/)

You can install the released version of vetiver from [PyPI](https://pypi.org/project/vetiver/):

```{python}
pip install vetiver
```

And the development version from [GitHub](https://github.com/tidymodels/vetiver-python) with:

```{python}
python -m pip install git+https://github.com/tidymodels/vetiver-python
```

## R

You can use vetiver with:

- a [tidymodels](https://www.tidymodels.org/) workflow
- [caret](https://topepo.github.io/caret/)
- [mlr3](https://mlr3.mlr-org.com/)
- [XGBoost](https://xgboost.readthedocs.io/en/latest/R-package/)
- [ranger](https://cran.r-project.org/package=ranger)
- [`lm()`](https://stat.ethz.ch/R-manual/R-patched/library/stats/html/lm.html) and [`glm()`](https://stat.ethz.ch/R-manual/R-patched/library/stats/html/glm.html)

You can install the released version of vetiver from [CRAN](https://cran.r-project.org/package=vetiver):

```{r}
install.packages("vetiver")
```

And the development version from [GitHub](https://github.com/tidymodels/vetiver-r) with:

```{r}
# install.packages("devtools")
devtools::install_github("tidymodels/vetiver-r")
```

:::
