---
title: "Get started with vetiver"
format:
  html:
    toc: true
    include-after-body:
      - sync-tabs.html
---

The vetiver framework for MLOps tasks is built for data science teams using R and/or Python, with a native, fluent experience for both. It is built to be extensible, with methods that can support many kinds of models.

## Installation

::: panel-tabset

## Python

You can use vetiver with:

- [scikit-learn](https://scikit-learn.org/)
- [PyTorch](https://pytorch.org/)

You can install the released version of vetiver from [PyPI](https://pypi.org/project/vetiver/):

```{python}
#| eval: false
pip install vetiver
```

And the development version from [GitHub](https://github.com/tidymodels/vetiver-python) with:

```{python}
#| eval: false
python -m pip install git+https://github.com/tidymodels/vetiver-python
```

## R

You can use vetiver with:

- a [tidymodels](https://www.tidymodels.org/) workflow
- [caret](https://topepo.github.io/caret/)
- [mlr3](https://mlr3.mlr-org.com/)
- [XGBoost](https://xgboost.readthedocs.io/en/latest/R-package/)
- [ranger](https://cran.r-project.org/package=ranger)
- [`lm()`](https://stat.ethz.ch/R-manual/R-patched/library/stats/html/lm.html) and [`glm()`](https://stat.ethz.ch/R-manual/R-patched/library/stats/html/glm.html)

You can install the released version of vetiver from [CRAN](https://cran.r-project.org/package=vetiver):

```{r}
#| eval: false
install.packages("vetiver")
```

And the development version from [GitHub](https://github.com/tidymodels/vetiver-r) with:

```{r}
#| eval: false
# install.packages("devtools")
devtools::install_github("tidymodels/vetiver-r")
```

:::

## Create a vetiver model

For this example, let's work with data on [hotel bookings](https://www.tidymodels.org/start/case-study/) to predict which hotel stays included children and which did not.

::: panel-tabset

## R

Let’s consider one kind of model supported by vetiver, a [tidymodels](https://www.tidymodels.org/) workflow that encompasses both feature engineering and model estimation.

```{r message = FALSE}
library(tidyverse)
library(tidymodels)

hotels <- read_csv('https://tidymodels.org/start/case-study/hotels.csv')

set.seed(123)
hotel_split <- initial_split(hotels, strata = children)
hotel_train <- training(hotel_split)
hotel_test  <- testing(hotel_split)

rf_recipe <- 
  recipe(children ~ ., data = hotel_train) %>% 
  step_date(arrival_date) %>% 
  step_holiday(arrival_date, keep_original_cols = FALSE)
rf_spec <- rand_forest(mode = "classification")

set.seed(234)
rf_fit <-
    workflow(rf_recipe, rf_spec) %>%
    fit(sample_frac(hotel_train, 0.5))
```


## Python

```{python}
1 + 1
```

:::

This `rf_fit` object is a fitted model, with both feature engineering and model parameters estimated using some of the training data `hotel_train`. We can create a `vetiver_model()` in R or `VetiverModel()` in Python from the trained model; a vetiver model object collects the information needed to store, version, and deploy a trained model.

::: panel-tabset

## R 

```{r}
#| message: false
library(vetiver)
v <- vetiver_model(rf_fit, "hotel_rf")
v
```


## Python

```{python}
1 + 1
```

:::

Think of this vetiver model as a deployable model object.

